Mendelian randomization for causal inference in genome-wide association studies

### 2022-03-02 mr-review.Rmd
BayesMR text added

Genome-wide association studies (GWAS) present unique opportunities for 
discovery of causal relationships via MR. A GWAS
interrogates millions of single nucleotide polymorphisms (SNPs) to infer which affect the trait [@the_wellcome_trust_case_control_consortium_genome-wide_2007;@klein_complement_2005].
In the nearly 17 years since publication of early studies, researchers have reported 
thousands of novel SNP-trait associations from GWAS [@klein_complement_2005;@welter_nhgri_2014;@visscher_10_2017;@visscher_five_2012]. Ever larger sample sizes in GWAS have enhanced statistical power to detect associations and refined our understanding of 
human health and disease.

@@

@@

>>>>

Genome-wide association studies (GWAS) present unique opportunities for 
discovery of causal relationships via MR [@swerdlow_selecting_2016]. A GWAS
interrogates millions of single nucleotide polymorphisms (SNPs) to infer which affect the trait [@the_wellcome_trust_case_control_consortium_genome-wide_2007;@klein_complement_2005].
In the nearly 17 years since publication of early studies, researchers have reported 
thousands of novel SNP-trait associations from GWAS [@klein_complement_2005;@welter_nhgri_2014;@visscher_10_2017;@visscher_five_2012]. Ever larger sample sizes in GWAS have enhanced statistical power to detect associations and refined our understanding of 
human health and disease.

@@

@hemani_automating_2017

@@

@bucur_inferring_2020 developed BayesMR, a Bayesian framework for two-sample MR with individual-level data and discuss an approximation for use with  
GWAS summary statistics. Unlike many MR methods, BayesMR both aims to control for 
horizontal pleiotropy and to quantitatively assess the possibility of reverse 
causation, where the putative exposure is actually caused by the putative outcome. 
For $J$ independent SNP instruments, @bucur_inferring_2020 permit each SNP instrument to 
have both a direct effect on the exposure and a horizontally pleiotropic direct effect 
on the outcome. Instrument effects and horizontally pleiotropic effects are assumed 
independent. Note that this is the Bayesian analog of the InSIDE assumption. 
After specifying the model, they present a nested sampling scheme [@skilling2006nested;@handley2015polychord1;@handley2015polychord] that simultaneously 
computes the model evidence for both causal directions (*i.e.*, exposure causes 
outcome and outcome causes exposure) and acquires samples from the posterior 
distribution for Bayesian statistical inferences.
